Reasoning About a Machine with Local Capabilities

### 2019-02-10 toplas/paper.tex
[TOPLAS] Added short paragraph about secure compilation

\citet{swasey:2017} have recently developed a \emph{logic}, OCPL, for verification
of object capability patterns. The logic is based on
Iris~\citep{iris,iris2,iris3}, a state of the art higher-order concurrent
separation logic and is formalized in Coq, building on the Iris Proof Mode for
Coq~\citep{ipm}. OCPL gives a more abstract and modular way of proving
capability safety for a lambda-calculus (with concurrency) compared to the
earlier work by \citet{Devriese:2016ObjCap}.
% domi: rest of the paragraph could be cut?
\begin{toplas}
In the future, we would also like to investigate a new program logic for
reasoning about capability safety for our capability machine model. In fact,
we think the lemmas in Section~\ref{sec:reasoning} are suggestive of the style
of results that could be captured in such a logic. We think Iris would also be
a natural starting point for such an endeavour since Iris is a
framework that can be instantiated with different programming languages. OCPL
was able to leverage existing Iris specifications for a high-level language;
for our capability machine model, however, it would be necessary to devise new
kinds of specifications for our low-level programs with unstructured
control-flow. It is likely that we could get inspiration from earlier work on
logics for assembly programming languages, such as XCAP~\citep{xcap}. Building a logic around the semantic model presented here would remove some of the tedious repetitive proof details making it more realistic to prove properties about larger more realistic programs.
\end{toplas}
\itoplassug{We could further add that this is not the level of abstraction that one would like to reason. One would like to reason about the programs we actually write - that is programs written in high-level languages. If we had a fully-abstract compiler from a nice high-level language to this low-level machine, then we could reason about the program in the high-level language and then because the compiler is fully-abstract, we would retain all the nice properties from the high-level language. How does this work fit in to this picture? The calling convention ensures two properties that we expect our nice high-level languages to have, namely well-bracketedness and local-state encapsulation. If we want to do fully-abstract compilation, then we need some way to guarantee this after compilation. Further, logical relations are often used in full-abstractness proofs. This work expands our knowledge about logical relations should be defined for low-languages and in particular capability machines. (maybe also mention that enforcement necessary for fully-abstract compilation. Could be capability, but it could also be something else - we just need something which we do not have now.)  }

>>>>

\citet{swasey:2017} have recently developed a \emph{logic}, OCPL, for verification
of object capability patterns. The logic is based on
Iris~\citep{iris,iris2,iris3}, a state of the art higher-order concurrent
separation logic and is formalized in Coq, building on the Iris Proof Mode for
Coq~\citep{ipm}. OCPL gives a more abstract and modular way of proving
capability safety for a lambda-calculus (with concurrency) compared to the
earlier work by \citet{Devriese:2016ObjCap}.
% domi: rest of the paragraph could be cut?
\begin{toplas}
In the future, we would also like to investigate a new program logic for
reasoning about capability safety for our capability machine model. In fact,
we think the lemmas in Section~\ref{sec:reasoning} are suggestive of the style
of results that could be captured in such a logic. We think Iris would also be
a natural starting point for such an endeavour since Iris is a
framework that can be instantiated with different programming languages. OCPL
was able to leverage existing Iris specifications for a high-level language;
for our capability machine model, however, it would be necessary to devise new
kinds of specifications for our low-level programs with unstructured
control-flow. It is likely that we could get inspiration from earlier work on
logics for assembly programming languages, such as XCAP~\citep{xcap}. Building a logic around the semantic model presented here would remove some of the tedious repetitive proof details making it more realistic to prove properties about larger more realistic programs.

If we want to scale this approach even further, we would like to reason at a
higher level of abstraction namely at the level of a high-level language.
That is, we would rather construct a program logic for a high-level language, so
we can reason about the programs the language we actually write them in and at
the same time get the guarantees for the programs we are going to execute, i.e.\
the compiled programs. To achieve this, we would have to construct a secure compilation that preserves the security abstraction of the high-level language. Our logical relation would not be directly applicable in a secure compilation proof, the experiences made from constructing a logical relation for a capability machine can still be used.
\end{toplas}
\itoplassug{We could further add that this is not the level of abstraction that
  one would like to reason. One would like to reason about the programs we
  actually write - that is programs written in high-level languages. If we had a
  fully-abstract compiler from a nice high-level language to this low-level
  machine, then we could reason about the program in the high-level language and
  then because the compiler is fully-abstract, we would retain all the nice
  properties from the high-level language. How does this work fit in to this
  picture? The calling convention ensures two properties that we expect our nice
  high-level languages to have, namely well-bracketedness and local-state
  encapsulation. If we want to do fully-abstract compilation, then we need some
  way to guarantee this after compilation. Further, logical relations are often
  used in full-abstractness proofs. This work expands our knowledge about
  logical relations should be defined for low-languages and in particular
  capability machines. (maybe also mention that enforcement necessary for
  fully-abstract compilation. Could be capability, but it could also be
  something else - we just need something which we do not have now.) }
